% !TeX program = lualatex

\documentclass[parskip=half-,12pt,oneside,openany,a5paper,numbers=endperiod]{scrbook}
\pdfvariable suppressoptionalinfo \numexpr1+2+4+8+16+32+64+128+256+512\relax

\usepackage{fontspec}
\setmainfont{TeX Gyre Pagella}
\setsansfont{TeX Gyre Heros}
\usepackage[scheme=plain,fontset=fandol]{ctex}
\setCJKmainfont{Noto Serif CJK SC}
\setCJKsansfont{Noto Sans CJK SC}
\usepackage{microtype}

\usepackage{unicode-math}
\setmathfont{TeX Gyre Pagella Math}
\usepackage{amsmath}

\usepackage{rotating}

\usepackage[colorlinks,allcolors=black]{hyperref}

\begin{document}
\title{Meltsand}
\author{Runxi Yu}
\date{2024}
\maketitle
\tableofcontents

\chapter{Leaf}

I am\\
\hphantom{I am} fluctuating\\
\hphantom{I am fluctuating} between two extremes

one extremely sticky spongy\\
\hspace*{2em}with its childish desire to be seen\\
\hspace*{4em}to be heard\\
\hspace*{6em}to be loved\\
\hspace*{8em}to drain

one extremely dry and separated\\
\hspace*{2em}with its sublime illusion\\
\hspace*{4em}distancing itself from everyone\\
\hspace*{6em}distancing itself from the ``other''\\
\hspace*{8em}distancing itself from the ``self''

deconstruct\\
reconsolidate

\chapter{flicker}

i am a flickering lightbulb\\
willing to go out any moment\\
evaporating off my tungsten\\
they diffuse across the vacuum\\
and tarnish the inside of my glass

\chapter{sugar glass}

lungs filled with pyro i press my finger against\\
my\\
forehead\\
fall, dissipate\\
i become transparent, i become a glass grenade

i wish that a coat of sugar could\\
turn me into\\
sugar\\
glass

glass with harmless little fragments\\
sharp at a glance but blunt as you approach

\chapter{独角兽}
\begingroup\addtolength{\baselineskip}{0.75ex}

有时候，玻璃独角兽会碎\\
但是祂的本质还在

那只角撞碎在地上，\\
破碎支离，折射着烛光，\\
在血肉模糊的身体上\\
闪耀着无畏的寂静\\
分崩离析，那沉醉中的时间

一口气吹过，\\
蜡烛灭了，而碎渣仍闪耀着\\
空气虚掩着烟雾和废墟\\
唤醒那种沉溺，\\
又浇上些许棉花

有时候，玻璃独角兽会碎；\\
但是祂的本质还在。

\endgroup

\chapter{两滴泪水}
\begingroup\addtolength{\baselineskip}{0.75ex}

两滴泪水，\\
一滴流下我的眼眶，\\
缓慢腐蚀着脸庞。\\
使疼痛开始变得粗糙，\\
细感\\
渗透，氧化。\\

两滴泪水，\\
一滴浸润我的心灵，\\
彻底席卷了大地。\\
使痛苦开始变得卑微，\\
愕然\\
却没有还原。

\endgroup

\chapter{灰}
\begingroup\addtolength{\baselineskip}{0.75ex}

光明瞬息即成灰，\\
强寒黑暗摧心深。

星火闪耀，破碎支离。

\endgroup

\chapter{Smush}

Begone the shallow fears and\\
Behind the unveiled face\\
There exists a skeleton\\
Consisting of two hun-\\
dred something bones\\
Skull cracked and shredded\\
Pinkish tissue irritably hopping\\
Oozing out of the seams\\
As if someone had dumped\\
A jar of red watercolor\\
Mixed with fresh lubricant

Bubbles of thoughts and memories\\
And desire and living and haps\\
Slowly emerging from the cerebrospinal\\
Goo, forming the occasional grid that\\
Seems to resemble a plate, a bell jar\\
A jar that I could see through, a jar\\
With a sharp unicorn's horn sticked\\
Hastily on top of it with\\
Cheap glue and duct tape

And the decaying flesh\\
Fills the jar.\\
Rotten, with its lurid colors\\
Decomposed aimlessly,\\
shedding fragments onto the\\
glass and cotton jar wall

I carefully wrap the jar in an\\
opaque paper box\\
then another stainless steel box\\
then another matte wood box\\
then a gift wrapper\\
Shiny, but covered with soothing colors\\
I print a sticker in 12-point Century font:\\
\qquad ``Life''\\
and stick it onto the wrapper.

Here you go. Your gift.

\chapter{There is no love}

\begingroup
\fontsize{50}{60}\selectfont
\begin{align*}
	\nexists\,\text{love}
\end{align*}
\endgroup

\rotatebox[origin=c]{180}{%
	\begin{minipage}[t]{\linewidth}
		But actually, \(\exists\,\text{y}\in\{x\,|\,\text{love} \backslash x\}\).
	\end{minipage}
}

\rotatebox[origin=c]{1}{%
	\begin{minipage}[t]{\linewidth}
		Would your response to ``do you love me'' be the same as your response to this question?
	\end{minipage}
}

\rotatebox[origin=c]{5}{%
	\begin{minipage}[t]{\linewidth}
		\scriptsize\itshape Perhaps. I guess.
	\end{minipage}
}

\chapter{The multiplicity returns}

May the ashes of my heart drift gently into the abyss of memory,
and let that memory consolidate.
A bright dot embellishing that unicorn's horn.
Then that glimmer of light shall decay, vaporize, and taint the glass walls of my body.
Gently running a cloth under my skin,
I wipe away every atom.
Every, single, one.
I stare at the cloth,
with rainbows seeping into its fabric,
assimilating into my body yet again.

\chapter{Sinus beat}

When my last sinus cell fires,\\
I shall still be alive.

When the door shuts,\\
The lights switched off,\\
The candle goes out,\\
The unicorn shattered,\\
The tuplips shiver in the austere cold,

When my chest wall collapses,\\
When the last drop of blood is pumped through my aorta,\\
I shall still be alive

\chapter{The Onion}

I peeled the onion\\
A layer shredded\\
Sent to oblision\\
Yet I dreaded\\
The heart of\\
The warped\\
Onion

With tis stained magenta glass surface\\
Amalgamating with every droplet\\
Of tears\\
Into each word engraved in my sight

I dropped the onion

Destitute of hope and love and faith

I walked away

I turned back

I picked it up again.

\chapter{Words and ink}

It is, I suppose, a particular kind of catharsis. The kind that lets me grasp
my chest and squeeze the blood out of it and experience some sort of, well,
exhilarating pain. Pain that feels like it could transcend my physical
existence, my experience of being part of a society, and the limitations of
being a human being. It's a powerful form of escape, one that I fear when
sober, but cling on to when desperate.

\appendix

\chapter{You, Again}

{}\hfill\textsc{Adrienne Rich}

\microtypesetup{protrusion=false}

Some nights I think you want too much. From me. I didn't ask\\
to parse again your idioms of littered\\
parking lots your chain-linked crane-hung sites\\
limp once more your crime-scene-festooned streets\\
to buildings I used to live in.\quad Lose my nerve\\
at a wrong door on the wrong floor\\
in search of a time.\quad The precision of dream is not\\
such a privilege. I know those hallways tiled in patterns\\
of oriental rugs those accordion-pleated\\
elevator gates. Know by heart the chipped\\
edges on some of those tiles. You who require this\\
heart-squandering want me wandering you, craving\\
to press a doorbell hear a lock turn, a bolt slide back\\
---always too much, over and over back\\
to the old apartment, wrong again, the key maybe\\
left with a super in charge of the dream who will not be found

\microtypesetup{protrusion=true}


\end{document}
